Doppler Imaging Fun

### 2019-06-18 tex/gitlinks.tex
woot

\newcommand{\codelink}[1]{\href{https://github.com/user/repo/blob/01e5eaaac7f32570634b33ad7aed7769c338a0d0/tex/figures/#1.py}{\codeicon}\,\,}
\newcommand{\animlink}[1]{\href{https://github.com/user/repo/blob/01e5eaaac7f32570634b33ad7aed7769c338a0d0/tex/figures/#1.gif}{\animicon}\,\,}
\newcommand{\prooflink}[1]{\href{https://github.com/user/repo/blob/01e5eaaac7f32570634b33ad7aed7769c338a0d0/tex/proofs/#1.ipynb}{\raisebox{-0.1em}{\prooficon}}}

>>>>

\newcommand{\codelink}[1]{\href{https://github.com/user/repo/blob/3428daf248495c61773970e1703192fbe198cb85/tex/figures/#1.py}{\codeicon}\,\,}
\newcommand{\animlink}[1]{\href{https://github.com/user/repo/blob/3428daf248495c61773970e1703192fbe198cb85/tex/figures/#1.gif}{\animicon}\,\,}
\newcommand{\prooflink}[1]{\href{https://github.com/user/repo/blob/3428daf248495c61773970e1703192fbe198cb85/tex/proofs/#1.ipynb}{\raisebox{-0.1em}{\prooficon}}}

### 2019-06-18 tex/ms.tex
woot

        \mathrm{e}^{-2\pi i k \xi} \mathrm{d}\xi

@@

        \mathrm{e}^{2\pi i k \xi} \mathrm{d}k

@@

of $\hat{I}(k, x, y)$ by an amount $\mathrm{e}^{-2\pi i k \Delta\xi}$

@@

    \mathrm{e}^{2\pi i k \alpha(x, y)}\hat{I_0}(k, x, y)

@@

        \mathrm{e}^{2\pi i k \alpha(x, y)}

@@

\citep[Equation B11 in][]{Luger2019}.

@@

If we take the inverse Fourier transform of Equation~(\ref{eq:fourier:S}),
we arrive at an expression for the observed spectrum:
%
\begin{align}
    \label{eq:convolve:SUgly}
    S(\xi) 
    &=
    \mathcal{F}^{-1} \Bigg[\,\,\,
        \iint\limits_{\mathcal{S}(x, y)}
        \tilde{\bvec{p}}^\top(x, y)
        \mathrm{e}^{2\pi i k \alpha(x, y)}
        \mathrm{d}\mathcal{S}(x, y)
        \cdot
        \bvec{A_1}
        \bvec{\hat{a}}(k)
    \Bigg] \nonumber \\
    &=
    \mathcal{F}^{-1} \Bigg[\,\,\,
        \iint\limits_{\mathcal{S}(x, y)}
        \tilde{\bvec{p}}^\top(x, y)
        \mathrm{e}^{2\pi i k \alpha(x, y)}
        \mathrm{d}\mathcal{S}(x, y)
    \Bigg]
    \circ
    \bvec{A_1}
    \bvec{a}(\xi)
    \quad ,
\end{align}
%
where the second line follows from the convolution theorem, with
$\circ$ denoting the convolution operator over $\xi$ and 
$\bvec{a}(\xi) \equiv \mathcal{F}^{-1}\left[ \bvec{\hat{a}}(k) \right]$.
Since only the exponential term in the integrand depends on $k$, we may 
further simplify this expression:
%
\begin{align}
    \label{eq:convolve:SUgly}
    S(\xi) 
    &=    
    \iint\limits_{\mathcal{S}(x, y)}
    \tilde{\bvec{p}}^\top(x, y)
    \mathcal{F}^{-1} \Big[
        \mathrm{e}^{2\pi i k \alpha(x, y)}
    \Big]
    \mathrm{d}\mathcal{S}(x, y)
    \circ
    \bvec{A_1}
    \bvec{a}(\xi)
    \nonumber \\
    &=    
    \iint\limits_{\mathcal{S}(x, y)}
    \tilde{\bvec{p}}^\top(x, y)
    \delta\big(\xi - \alpha(x, y)\big)
    \mathrm{d}\mathcal{S}(x, y)
    \circ
    \bvec{A_1}
    \bvec{a}(\xi)
    \quad,
\end{align}
%
where $\delta$ is the Dirac delta function. Since the integrand in the
equation above is zero everywhere except where the argument of the
$\delta$ function is zero, the equation may be written as a line integral 
along some curve $\mathcal{C}(x, y)$ that satisfies $\alpha(x, y) = \xi$
and $x^2 + y^2 \le 1$:
%
\begin{align}
    \label{eq:convolve:SUgly}
    S(\xi) 
    &=    
    \int\limits_{\mathcal{C}(x, y)}
    \tilde{\bvec{p}}^\top(x, y)
    \mathrm{d}\mathcal{C}(x, y)
    \circ
    \bvec{A_1}
    \bvec{a}(\xi)
    \quad.
\end{align}
%

>>>>

        \mathrm{e}^{-2\pi \imag k \xi} \mathrm{d}\xi

@@

        \mathrm{e}^{2\pi \imag k \xi} \mathrm{d}k

@@

of $\hat{I}(k, x, y)$ by an amount $\mathrm{e}^{-2\pi \imag k \Delta\xi}$

@@

    \mathrm{e}^{2\pi \imag k \alpha(x, y)}\hat{I_0}(k, x, y)

@@

        \mathrm{e}^{2\pi \imag k \alpha(x, y)}

@@

\citep[Equation B11 in][]{Luger2019}. On the observer-facing surface of 
the unit sphere, the coordinate $z$ is given by $\sqrt{1 - x^2 - y^2}$.

@@

If we take the inverse Fourier transform of Equation~(\ref{eq:fourier:S}),
we arrive at an expression for the observed spectrum:
%
\begin{align}
    \label{eq:convolve:SUgly1}
    S(\xi) 
    &=
    \mathcal{F}^{-1} \Bigg[\,\,\,
        \iint\limits_{\mathcal{S}(x, y)}
        \tilde{\bvec{p}}^\top(x, y)
        \mathrm{e}^{2\pi \imag k \alpha(x, y)}
        \mathrm{d}\mathcal{S}(x, y)
        \cdot
        \bvec{A_1}
        \cdot
        \bvec{\hat{a}}(k)
    \Bigg] \nonumber \\
    &=
    \mathcal{F}^{-1} \Bigg[\,\,\,
        \iint\limits_{\mathcal{S}(x, y)}
        \tilde{\bvec{p}}^\top(x, y)
        \mathrm{e}^{2\pi \imag k \alpha(x, y)}
        \mathrm{d}\mathcal{S}(x, y)
        \cdot
        \bvec{A_1}
    \Bigg]
    \circ
    \bvec{a}(\xi)
    \quad ,
\end{align}
%
where the second line follows from the convolution theorem, with
$\circ$ denoting the convolution operator over $\xi$ and 
$\bvec{a}(\xi) \equiv \mathcal{F}^{-1}\left[ \bvec{\hat{a}}(k) \right]$.
Since only the exponential term in the integrand depends on $k$, we may 
further simplify this expression:
%
\begin{align}
    \label{eq:convolve:SUgly2}
    S(\xi) 
    &=    
    \iint\limits_{\mathcal{S}(x, y)}
    \tilde{\bvec{p}}^\top(x, y)
    \mathcal{F}^{-1} \Big[
        \mathrm{e}^{2\pi \imag k \alpha(x, y)}
    \Big]
    \mathrm{d}\mathcal{S}(x, y)
    \cdot
    \bvec{A_1}
    \circ
    \bvec{a}(\xi)
    \nonumber \\
    &=    
    \iint\limits_{\mathcal{S}(x, y)}
    \tilde{\bvec{p}}^\top(x, y)
    \delta\big(\xi - \alpha(x, y)\big)
    \mathrm{d}\mathcal{S}(x, y)
    \cdot
    \bvec{A_1}
    \circ
    \bvec{a}(\xi)
    \quad,
\end{align}
%
where $\delta$ is the Dirac delta function. Since the integrand in the
equation above is zero everywhere except where the argument of the
$\delta$ function is zero, the equation may be written as a line integral 
along some curve $\mathcal{C}(x, y)$ whose points satisfy 
$\alpha(x, y) = \xi$
and lie within $\mathcal{S}$:
%
\begin{align}
    \label{eq:convolve:S}
    S(\xi) 
    &=    
    \mathbf{g}(\xi)
    \circ
    \bvec{a}(\xi)
    \quad,
\end{align}
%
where we define
%
\begin{align}
    \label{eq:convolve:g}
    \mathbf{g}(\xi) 
    &\equiv    
    \mathbf{s}^\top(\xi) 
    \cdot
    \bvec{A_1}
\end{align}
%
and
%
\begin{align}
    \label{eq:convolve:sT}
    \mathbf{s}^\top(\xi) 
    &\equiv    
    \int\limits_{\mathcal{C}(x, y)}
    \tilde{\bvec{p}}^\top(x, y)
    \mathrm{d}\mathcal{C}(x, y)
    \quad.
\end{align}

In order to solve Equation~(\ref{eq:convolve:S}), we must assume 
explicit functional
forms for the velocity field $\alpha(x, y)$ and the region of integration
$\mathcal{S}$, which we do in the following sections.

\section{Doppler imaging of rigid rotators}
\label{sec:rigid}
%
In the case that the star's rotation is rigid (i.e., differential rotation
is negligible), the radial velocity at any point on the surface is 
simply proportional to the distance to the axis of rotation. Without loss
of generality, if we assume the axis of rotation lies along the $y-z$ plane,
we have
%
\begin{align}
    \beta(x, y) = \frac{\omega \sin i \, x}{c}
\end{align}
%
and
%
\begin{align}
    \label{eq:rigid:alpha}
    \alpha(x, y) = \ln\left( 1 - \frac{\omega \sin i \, x}{c} \right)
    \quad ,
\end{align}
%
where $\omega$ is the angular velocity of the star and $i$ is the
inclination relative to $\hat{y}$. Furthermore, if
we assume the star is unocculted, the region of integration $\mathcal{S}$ 
is simply the unit disk, so $\mathcal{C}(x, y)$ is the set of all points
that satisfy both $\alpha(x, y) = \xi$ and $x^2 + y^2 \le 1$.
Solving Equation~(\ref{eq:rigid:alpha}) for $x$ and 
setting $\alpha(x, y) = \xi$, we find that the curves $\mathcal{C}(x, y)$ 
are simply vertical lines on the unit disk given by
%
\begin{align}
    x &= 
        \frac{c}{\omega\sin i} \left(1 - \mathrm{e}^{\xi}\right)
    \nonumber \\
    &\equiv \gamma(\xi)
    \quad .
\end{align}
%
The integral in Equation~(\ref{eq:convolve:sT}) is therefore just an integral
over $y$:
%
\begin{align}
    \label{eq:rigid:sT}
    \bvec{s}^\top(\xi) 
    &=    
    \int\limits_{-\sqrt{1 - \gamma(\xi)^2}}^{\sqrt{1 - \gamma(\xi)^2}}
    \tilde{\bvec{p}}^\top
    \Big(\gamma(\xi), y\Big)
    \mathrm{d}y
    \quad .
\end{align}
%
From \citet{Luger2019}, the $n^\mathrm{th}$ term of the polynomial basis
is a polynomial in $x$, $y$, and $z = \sqrt{1 - x^2 - y^2}$:
%
\begin{proof}{PolynomialBasis}
    \tilde{p}_n 
    &=
    x^i y^j z^k
\end{proof}
%
where $i, j, k$ are integers given by
%
\begin{proof}{PolynomialBasis}
    \label{eq:lm}
    i &= \floor*{\Lambda - \Delta}
    \nonumber \\[0.5em]
    j &= \floor*{\Delta}
    \nonumber \\[0.5em]
    k &= \ceil*{\Delta} - \floor*{\Delta}
\end{proof}
%
with
%
\begin{proof}{PolynomialBasis}
    \Lambda &= \floor*{\sqrt{n}}
    \nonumber \\[0.5em]
    \Delta &= \frac{n - \Lambda^2}{2}
    \quad .
\end{proof}
%
We may now explicitly write down the $n^\mathrm{th}$ term of 
Equation~(\ref{eq:rigid:sT}),
%
\begin{align}
    s_n(\xi) 
    &=    
    \begin{dcases}
        \gamma(\xi)^i
        \int\limits_{-\sqrt{1 - \gamma(\xi)^2}}^{\sqrt{1 - \gamma(\xi)^2}}
            y^j
        \mathrm{d}y
        &
        \qquad k = 0
        \\[1em]
        \gamma(\xi)^i
        \int\limits_{-\sqrt{1 - \gamma(\xi)^2}}^{\sqrt{1 - \gamma(\xi)^2}}
            y^j
            \sqrt{1 - \gamma(\xi)^2 - y^2} \,
        \mathrm{d}y 
        & 
        \qquad k = 1 \quad ,
    \end{dcases}
\end{align}
%
and solve it analytically:
%
\begin{align}
        \label{eq:rigid:sTAnalytic}
        s_n(\xi) 
        &=     
    \begin{dcases}
        \frac{2 \gamma(\xi)^i \big(1 - \gamma(\xi)^ 2\big)^{\frac{j + 1}{2}}}{j + 1}
        &
        \qquad k = 0 \,\,\, \& \,\,\, j \, \mathrm{even}
        \\[1em]
        \gamma(\xi)^{i} \mathcal{I}_{j}\big(\gamma(\xi)\big)
        &
        \qquad k = 1 \,\,\, \& \,\,\, j \, \mathrm{even}
        \\[1em]
        0
        &
        \qquad j \, \mathrm{odd} \quad,
    \end{dcases}
\end{align}
%
where $\mathcal{I}_j(\gamma)$ is computed analytically
from the upward recurrence relation given by Equation~(\ref{eq:app:Ij}),
with $\mathcal{I}_0(\gamma)$ given by Equation~(\ref{eq:app:I0}).

\appendix

\section{Integrals}
\subsection{Equation~(\ref{eq:rigid:sTAnalytic})}
%
When $k = 1$, we may write the $n^\mathrm{th}$ component
of Equation~(\ref{eq:rigid:sTAnalytic}) as
%
\begin{align}
    s_n(\xi) = 
        \gamma(\xi)^{i}
        \mathcal{I}_{j}\big(\gamma(\xi)\big)
\end{align}
%
where
%
\begin{align}
    \mathcal{I}_j(\gamma) &\equiv
    \int\limits_{-\sqrt{1-\gamma^2}}^{\sqrt{1-\gamma^2}}
        y^{j}
        \sqrt{1 - \gamma^2 - y^2} \,
    \mathrm{d}y
    \quad.
\end{align}
%
Substituting $y = \sin\psi\sqrt{1 - \gamma^2}$, we have
%
\begin{align}
    \mathcal{I}_j(\gamma) &=
    %
    (1 - \gamma^2)^{\frac{j + 2}{2}}
    \int\limits_{-\frac{\pi}{2}}^{\frac{\pi}{2}}
        \sin^j\psi
        \cos^2\psi \,
    \mathrm{d}\psi
    \quad.
\end{align}
%
We may solve this integral using the recurrence relation
%
\begin{align}
    \int
        \sin^j\psi
        \cos^m\psi \,
    \mathrm{d}\psi
    &=
    -\frac{\sin^{j-1}\psi \cos^{m+1}\psi}{j + m}
    +
    \frac{j - 1}{j + m}\int\sin^{j-2}\psi \cos^m\psi \mathrm{d}\psi
    \quad ,
\end{align}
%
which, in our case, simplifies to
%
\begin{align}
    \int\limits_{-\frac{\pi}{2}}^{\frac{\pi}{2}}
        \sin^j\psi
        \cos^2\psi \,
    \mathrm{d}\psi
    &=
    \frac{j - 1}{j + 2}\int\limits_{-\frac{\pi}{2}}^{\frac{\pi}{2}}\sin^{j-2}\psi \cos^2\psi \mathrm{d}\psi
    \quad.
\end{align}
%
We thus arrive at the recurrence relation
%
\begin{align}
    \label{eq:app:Ij}
    \mathcal{I}_j(\gamma) &=
    \frac{j - 1}{j + 2}
    (1 - \gamma^2)
    \mathcal{I}_{j-2}(\gamma)
    \quad.
\end{align}
%
Since this expression is only used when $j$ is even, we only need one
initial condition for upward recursion:
%
\begin{align}
    \label{eq:app:I0}
    \mathcal{I}_0(\gamma) &\equiv
    \int\limits_{-\sqrt{1-\gamma^2}}^{\sqrt{1-\gamma^2}}
        \sqrt{1 - \gamma^2 - y^2} \,
    \mathrm{d}y
    \nonumber \\
    &= \frac{\pi}{2}(1 - \gamma^2)
    \quad.
\end{align}
